```{r include=FALSE}
source("common.R") 
```

# Introduction {#intro}

You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].


# Introduction {#intro}
This chapter introduces the dataset that we will be using and cleaning the data for the next chapters in this books. 

## Dataset 
We will be using the _Wave 6_ of the [World Values Survey](https://www.worldvaluessurvey.org/WVSDocumentationWV6.jsp) for this project. The data for this survey was collected between 2010 and 2014. 
\n 
We will be using the following libraries for data cleaning 
```{r}
library(tidyverse)
library(reactable)
```


There are `r (ncol(data_raw))` columns in the data. We will be retaining variables which consist of self-report data scaling from 1 to 5 or 1 to 7. 

## Functions
Some functions involving outputting data and workflow operations being used in the global environment will be introduced here. I will provide explanation about each function in the comments. 
```{r}
#Function 1
  #Call and pring the value for a specific country 

 cntry_n <- function(country_code) {
   data_raw %>% 
     count(C_COW_ALPHA) %>% 
     filter(C_COW_ALPHA == {{country_code}}) 
 }

#Function 2 
  #ncol and nrow at the same time

length_CnR <- function(data){  
  n_col <- ncol(data)
  n_row <- nrow(data)
  return(tibble(n_col, n_row))
}
```

## Data 
\n
We will retain data for three countries for this analysis. 
Three countries considered to be least similar on aspects such as language, social structures and values will be retained. Let's look at the sample sizes for each country. 

```{r}
data_raw %>% 
  count(C_COW_ALPHA) %>% 
  arrange(desc(n)) %>% 
  
  reactable(
    defaultColDef = colDef(
    cell = function(value) format(value, nsmall = 1),
    align = "center",
    minWidth = 70,
    headerStyle = list(background = "#f7f7f8")
    ),
  
    
    columns = list(
    C_COW_ALPHA = colDef(name = "Country Code"),
    n = colDef(name = "Sample size (n)")
    ),
    
  bordered = TRUE,
  highlight = TRUE
  )
```

\n
Three countries for this analysis include - The United States of America (USA) ( _n_ = `r (cntry_n("USA")$n)`), India (IND) ( _n_ = `r (cntry_n("IND")$n)`) and Nigeria (NIG) ( _n_ = `r (cntry_n("NIG")$n)`)
\n
We will retain data for these three countries with some data cleaning. We will remove data of those participants who have 20% or more missing values. 
```{r}
 data = data_raw %>% 
   filter(C_COW_ALPHA %in% c("USA","IND","NIG")) %>% 
  rename("country" = C_COW_ALPHA) %>% 
  dplyr::select(country, V4:V9, V12:V22, V45:V56, V70:V79, V95:V146, -V144, V153:V160J, V192:V216, V228A:V228K) %>% 
  #remove cols which have all NA's
  janitor::remove_empty(which = "cols") %>% 
  #remove cols which have 20% or more missing data (NA)
  purrr::discard(~sum(is.na(.x))/length(.x)* 100 >=20) %>% 
  #impute the missing values with 
  mutate_all(~ifelse(is.na(.x), mean(.x, na.rm = TRUE), .x)) %>% 
  remove_rownames()
```

Here is a the same data transposed, where rows are variables and columns are participants. 
```{r}
#Transpose data 
data_t <- data %>% 
  mutate(ind = paste0("p", 1:nrow(data))) %>% 
   dplyr::select(ind, everything(), -country) %>% 
  t %>% as.data.frame %>% janitor::row_to_names(1) %>% 
  mutate_all(funs(as.numeric(as.character(.))))
```



