# Introduction {#intro}

You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].


Calling data
```{r}
# ----Functions - Data wrangling fro congruence coefficient

#Call data = Ipsatized or Original data 
files <- function(dataType){
  dir_ls(here::here("data"))%>%
  as_tibble() %>% 
  filter(str_detect(value, {{dataType}}))
}
```
## Check here now 
Some more code doing something else

```{r}
#creating the main col names - factor# and dpecify g or ng
rename_df <- function(d) {
  #look for the col which has sum of NA's equal to the number of rows in the data frame - we want to remove this blank col - output is logical
  col_miss_log <- apply(d, 2, function(x) sum(is.na(x)) == nrow(d))
  #Give the col which has logical = T; this is the blank col
  col_miss <- seq_along(d)[col_miss_log]
  #Remove the first col, the blank col and the co next to blank col
  d_subset <- d[ ,-c(1, col_miss, col_miss + 1)]
  #create the main stem for g_
  nms1 <- gsub("^(\\d?\\d).+", "g_\\1", names(d_subset)[seq(1, col_miss - 2)])
  nms2 <- gsub("^(\\d?\\d).+", "ng_\\1", names(d_subset)[seq(col_miss - 1, length(names(d_subset)))])
  names(d_subset) <- c(nms1, nms2)
  d_subset
}

```


